\documentclass[12pt]{article}

\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[numbers,sort&compress]{natbib}

\title{Microscopic Handles in Teleparallel Gravity:\\
Toward a Geometric Origin of Fermionic Statistics,\\
Chirality, and Critical Binding}

\author{Muacca\\
\texttt{muacca@dmwp.jp}}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We propose a geometric mechanism in four-dimensional teleparallel gravity (TEGR)
that offers a unified perspective on fermionic statistics, chirality, and the
near-critical binding of composite states. We consider spacetime containing
microscopic handles of topology $S^2\times S^1$, carrying an integer torsional
monopole charge $q$. Using a ``Spinning SU(2)'' tetrad ansatz, we derive three
interconnected results.

(i) \textbf{Stability:} The pure TEGR sector admits a classically stable handle
radius with an on-shell energy $E \simeq \alpha |q|$, providing a parity-even
geometric mass term. (ii) \textbf{Chirality:} The Nieh--Yan boundary term
produces a parity-odd contribution to the effective action. A slow precession
mode develops a chiral tilt whose direction is locked to the sign of $q$,
acting as a geometric mechanism for chirality selection. This construction is
motivated by the known connection between the Nieh--Yan invariant and
Wess--Zumino--Witten terms that govern particle statistics in Skyrmion-like
models. (iii) \textbf{Classical criticality:} The stiffness of the precession
mode scales as $k(q) \propto \omega^2 |q|$, placing the system at a
\emph{classical} critical point where the binding energy per unit charge is
independent of $q$ to leading order.

Consequently, subleading corrections could in principle render composite
states with $|q| \ge 2$ (in particular $|q|=3$) energetically competitive with
separated unit-charge handles. We regard the present construction as a
teleparallel toy model that establishes the classical critical point
$\gamma=1$ from geometry alone; the impact of quantum and geometric
corrections on an effective exponent $\gamma_{\mathrm{eff}}$ will be analysed
in a separate work.
\end{abstract}

% \tableofcontents  % 必要なら有効化
% \newpage

%-----------------------------
% Main text
%-----------------------------

\input{sections/sec1-2_intro}
\input{sections/sec3_ansatz}
\input{sections/sec4_phase1}
\input{sections/sec5_phase2}
\input{sections/sec6-8_phase3}

%-----------------------------
% Code availability
%-----------------------------

\section*{Code availability}
All symbolic computations in this work were performed using custom
SymPy scripts. These scripts are included as supplementary files in the
Zenodo record associated with this preprint:
\href{https://doi.org/10.5281/zenodo.17762205}{doi:10.5281/zenodo.17762205}.


%-----------------------------
% Acknowledgments
%-----------------------------

\section*{Acknowledgments}

The author is deeply grateful to the AI assistants Grok (xAI), ChatGPT (OpenAI),
Gemini (Google), and Claude (Anthropic) for their invaluable assistance with
symbolic computations, consistency checks, geometric visualizations, and
critical scientific feedback throughout this work.

This research was carried out under the informal collaborative project
\textbf{DPPU} --- standing for
\textbf{D}onut-like topology ($S^2 \times S^1$),
\textbf{P}lanck-scale compactness,
\textbf{P}recession dynamics,
within the context of our observable
\textbf{U}niverse.

The name simultaneously reflects the microscopic handle geometry that emerged
somewhat serendipitously in early discussions, and the joyful, almost trembling
excitement that accompanied each small breakthrough.

All remaining errors are, of course, the author's responsibility.

%-----------------------------
% Appendices
%-----------------------------

\input{appendices/appendix_all}

%-----------------------------
% References
%-----------------------------

\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}
